Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

## Measurement
In the current presidential election cycle,  Americans’ opinions on voting for their preferred candidate are transformed into data points through a series of steps that involve surveying, data processing, and structuring responses. Pollsters then store these responses in structured databases for analysis. In the dataset obtained from Project 538, each entry in the dataset corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for each candidate. Each pollster has different approaches to sampling respondents, recruiting respondents, as well as questionnaire types, which affects each pollster’s reliability and data quality.

@@

# Pollster Methodology Overview and Evaluation
The following information about the Siena College/New York Times poll methodology is based on the details provided in the article "How The Times/Siena Poll Is Conducted" by the New York Times [@nyt-siena-methodology].

The Siena College/New York Times poll for the 2024 presidential election uses a robust sampling approach to ensure accuracy and relevance. The poll applies a stratified dual-frame sample, drawing from both land-lines and cell phones. Each poll is conducted by phone using live interviewers at call centers based in Florida, New York, South Carolina, Texas, and Virginia. The sampling population is all registered voters who live in the six battleground states: Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin.

Given the nature of the presidential elections, the decision is based on the electoral college instead of the popular vote, thus the pollster focuses on polling on the states that are likeliest to decide the outcome of the race. The voters’ information is taken from the L-2 voter file, which includes details such as voter registration and history of participation in previous elections.

To refine their sample, Siena College adjusts the data by key demographic and political variables such as region, race/ethnicity, party affiliation, education, and voting patterns from the 2020 election. These adjustments help ensure the sample accurately represents the electorate. Additionally, they use a “likely voter screen,” which combines self-reported likelihood of voting with historical voter behavior to estimate how likely respondents are to vote in the upcoming election.

At every step of the survey, Sienna/NYT uses the information in the data to try to ensure that they have the right number of Democrats and Republicans, young people and old people, the right ratio of people with different income levels, and a diverse mix of different races and regions. Once the survey is complete, they compare their respondents to the voter file and use a process known as weighting to ensure that the sample reflects the broader voting population. This combination of historical data and weighted adjustments ensures the poll is designed to predict election outcomes as accurately as possible.

The New York Times/Siena College pollsters handle non-response by using weighting and adjusting their sample to correct for any biases that might emerge due to individuals not responding. Specifically, they adjust on multiple demographic and political variables such as age, gender, education, race/ethnicity, and party affiliation. They also account for variations in voter likelihood and previous voting patterns, ensuring that the sample represents the likely electorate as accurately as possible.

This helps reduce the potential bias from non-response, especially since certain demographic groups or political affiliations may be less likely to respond to polls. By rebalancing the sample, they ensure that even if some groups have lower response rates, their representation in the final poll results aligns with what is expected based on historical trends and current voter enthusiasm.

# Idealized Methodology

Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions. This approach incorporates best practices and innovative techniques to ensure a comprehensive prediction of voter behavior in the upcoming 2024 U.S. presidential election. To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@US-election-survey] under [@sec-references].

>>>>

## Measurement
In the current presidential election cycle,  Americans’ opinions on voting for their preferred candidate are transformed into data points through a series of steps that involve surveying, data processing, and structuring responses. Pollsters then store these responses in structured databases for analysis. In the dataset obtained from Project 538 [@FiveThirtyEight], each entry in the dataset corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for each candidate. Each pollster has different approaches to sampling respondents, recruiting respondents, as well as questionnaire types, which affects each pollster’s reliability and data quality.

@@

# Pollster methodology overview and evaluation: NYT/Siena College

## Background of NYT/Siena College Polling
The NYT/Siena College Polling is a U.S. polling organization where the New York Times and Siena College collaborate to deliver precise and timely poll results for the presidential elections [@nyt-siena-methodology]. For the 2024 election cycle, The NYT/Siena College poll uses a robust sampling approach to ensure accuracy and representativeness. The poll applies a stratified dual-frame sample, drawing from both landlines and cell phones.

## Target population, Sampling frame, and Sample
- Target population: Target population refers to “the collection of all items about which we would like to speak” [@telling-stories-data]. In this case, NYT/Siena College’s polling target population is all registered American voters who live in the six battleground states: Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin. Given the nature of the presidential elections, the election result is based on the electoral college instead of the popular vote, thus the pollster focuses on polling on the states that are the likeliest to decide the outcome of the race, which are called battleground states. The battleground states are states that have similar voter support for both the Democratic or the Republican party, a slight difference in support could determine which party wins a state.
- Frame: Sampling frame refers to “a list of all items from the target population that we could get data about” [@telling-stories-data]. NYT/Siena College’s sampling frame is a comprehensive list of registered American voters obtained from the L-2 voter file, it includes details such as demographic characteristics, contact information, and voting history. By using the L-2 voter file as their sampling frame, the pollster can ensure representativeness by allowing them to draw samples that accurately reflect the demographics of the registered voter population. 
- Samples: Sample refers to “the items from the sampling frame that we get data about” [@telling-stories-data]. In this case, the samples of NYT/Sienna College are individuals who are registered American voters obtained from the L-2 voter file from each stratum from each state who have also answered and completed the questionnaires.

## Sampling Methodology
NYT/Siena College uses stratified sampling as their methodology where the population is divided into distinct strata based on shared characteristics. These strata include age, gender, race, geographic region, and other various relevant variables. Once the population is divided, random samples are taken from each stratum, in proportion to their size in the population. For each poll, NYT/Siena College takes a sample size of 1000 registered American voters. To refine their sample, NYT/Siena College adjusts the data by key demographic and political variables such as region, race/ethnicity, party affiliation, education, and voting patterns from the 2020 election. These adjustments help ensure the sample accurately represents the electorate. Additionally, they use a “likely voter screen,” which combines self-reported likelihood of voting with historical voter behavior to estimate how likely respondents are to vote in the upcoming election. At every step of the survey, Siena/NYT uses the information in the datasets to try to ensure that they have the right number of Democrats and Republicans, young people and old people, the right ratio of people with different income levels, and a diverse mix of different races and regions. Once the survey is complete, they compare their respondents to the voter file and use a process known as weighting to ensure that the sample reflects the broader voting population. This combination of historical data and weighted adjustments ensures the poll is designed to predict election outcomes as accurately as possible.

## Trade-off of Sampling Methodology
Although stratified sampling has many strengths such as appropriate representation of the target population, reduction of sampling bias, and sampling efficiency, it also has some weaknesses. First, its implementation can be complex, stratified sampling requires detailed knowledge of the population to define the strata appropriately. This can be challenging to implement if accurate, up-to-date information on demographic or geographic characteristics is unavailable or difficult to access. For example, if there are significant changes in the L-2 voter file since the last election, it is difficult to take account into the unknown changes.
Second, there could be potential for mis-stratification. If the chosen strata do not capture meaningful differences within the population, or if important subgroups are overlooked, the results might be biased or inaccurate. For example, if new voting trends emerge that were not captured in the strata definitions, the poll might miss important shifts in voter opinion.
Third, stratified sampling can be time-consuming and resource-intensive. Diving the population into strata and ensuring proportionate sampling within each group can be more time-consuming and resource-intensive compared to simple random sampling. Pollsters may need to gather detailed data on the target population, which adds to the cost and complexity of the survey. 
Lastly, there is the risk of over-emphasis on stratification. If the population is stratified too finely, the sample size in each stratum may become too small to yield meaningful results, increasing the margin of error for each subgroup.

## Sample Recruitment Methods
Each poll is conducted by phone using live interviewers at call centers based in Florida, New York, South Carolina, Texas, and Virginia. Most people that the live interviewers call do not answer the call, and some don’t finish the survey, leaving only about two percent both answer their phone and complete the poll. To mitigate this:
First, the pollster keeps calling numbers until someone picks up, even though it can take hours–a reason why a high-quality poll is time-consuming and expensive to conduct. 
Second, voter files are used to help make sure the pollster is reaching a representative sample; for example, they may reach out to a higher percentage of people who are less likely to respond. 
Third, pollster mathematically adjusts the responses they do receive so they are representative of the wider population, it includes adjusting by age, race, gender, voting history and etc.

## Non-response Handling
Non-responses are critical factors in polling because they affect the representativeness of the sample and, consequently, the accuracy of the poll results. When a significant portion of selected individuals doesn’t respond, the poll risks becoming unrepresentative of the broader population, especially if certain groups are systematically more or less likely to respond. It is also important to know how pollsters handle non-response since it provides transparency about the polling methodology. It also gives confidence that the results are not skewed by an unrepresentative sample when pollsters clearly explain their non-response handling. Therefore, polls with thorough non-response strategies are generally more reliable than those without. The NYT/Siena College pollsters handle non-response by using weighting and adjusting their sample to correct for any biases that might emerge due to individuals not responding. Specifically, they adjust on multiple demographic and political variables such as age, gender, education, race/ethnicity, and party affiliation. They also account for variations in voter likelihood and previous voting patterns, ensuring that the sample represents the likely electorate as accurately as possible. This helps reduce the potential bias from non-response, especially since certain demographic groups or political affiliations may be less likely to respond to polls. By rebalancing the sample, they ensure that even if some groups have lower response rates, their representation in the final poll results aligns with what is expected based on historical trends and current voter enthusiasm.

## Questionnaire Pros and Cons
Based on the NYT/Siena College’s past practices on presidential election polls, they often ask questions about candidate preference, approval ratings, voter opinions on key issues, economic perceptions, as well as social and cultural issues. They also ask about the respondents’ demographic information and voting behavior to categorize respondents for analysis.

### Strengths of the NYT/Siena College questionnaires
First, they have a comprehensive question design. NYT/Sienna College polls ask a wide range of questions that provides a detailed view of the respondent’s preferences and opinions. It helps produce multi-dimensional data that allows for in-depth analysis and segmentation of voter groups. 
Second, the questionnaires are tailored to likely voters, as identified through the L-2 voter file. This means that the voters are more likely to be up-to-date with relevant policies proposed by different parties, and have formed their own opinions about them, this improves the predictive reliability of the poll. 
Third, the questionnaires contain detailed demographic questions about the respondents which allow NYT/Siena College to stratify responses by voter subgroups, identifying varied voter bases and diverse issues.

### Weaknesses of the NYT/Siena College questionnaires
First, the length and complexity of the questionnaires might cause the respondents to be less likely to respond or finish the polls. Respondents may rush through questions or drop out before completion. Respondents may also be prone to moderacy response bias where they have the tendency to choose middle options regardless of question content, as well as response order bias where respondents choose answers based on their order. 
Second, questionnaires have a dependence on self-reported likelihood to vote. Self-reported likelihood can be unreliable, especially for individuals whose intentions to vote may change close to the actual election date. This can reduce the accuracy of predictions based on likely voter models. 
Third, there are limited open-ended questions. Many voters may have mixed opinions on specific issues or policies that cannot be explained in a single-choice answer. However, answers to open-ended questions are often difficult to convert to actionable data, so questionnaires tend to rely on closed-ended questions for scalability. This might restrict the ability to capture complex reasonings behind their choices, which could add extra strain on the time and resources used to interpret the answers.

# Idealized Methodology
Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions. This approach incorporates best practices and innovative techniques to ensure a comprehensive prediction of voter behavior in the upcoming 2024 U.S. presidential election. To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@US-election-survey] under [@sec-references].
